# DaSEH Project

## Data Used

I am using USGS pesticide estimate data, Census Bureau data on U.S. county boundaries, and historical heat records in the U.S. from X (). 

USGS Estimated Annual Agricultural Pesticide Use for Counties of the Conterminous United States, 2013-2017 (ver. 2.0, May 2020) data were obtained from here: https://www.sciencebase.gov/catalog/item/5e95c12282ce172707f2524e. Cartographic Boundary Files from the U.S Census Bureau were obtained from here for the year 2013: https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.2013.html#list-tab-1556094155.



```{r}

```

**describe** what your data looks like and **where** you got it. Identify what the variables and samples are, or point to a data dictionary. Describe how the data was originally created. 


# Data Cleaning/Wrangling

Perform any data subsetting, cleaning, or manipulation methods that were described in this course on your data. Examples are: renaming the columns, recoding values, reshaping the data, filtering the data etc. 

You might also research other methods to include here, especially if your data is very unique.

```{r}

```


# Data Visualization

Create some visualizations of your data using the esquisse app or the ggplot2 package.

```{r}

```


# Data Analysis

Perform a **simple analysis** of your data. This can involve summarizing the data to describe aspects about it (quartiles, means, range etc.) or a simple statistical test. 

Don't forget to describe what analysis you performed and why. Provide some simple **interpretation** about what your analysis might indicate about your data.

```{r}

```


# Please include additional sections as desired.

```{r}

```


# Versions

Include some information about the version of R and the packages you are using by running sessionInfo().

```{r}
sessionInfo()
```

Render your document to an html or pdf using the "Knit" button.
