---
title: "DaSEH Lightning Talk"
author: "Cecilia Martindale"
date: "2024-07-31"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Load packages, installing if needed
if (!requireNamespace("pacman", quietly = TRUE))
  install.packages("pacman")
pacman::p_load(
  RColorBrewer,
  tidyr,
  tidyverse,
  ggplot2,
  MetBrewer,
  stringr,
  ggpubr,
  here,
  sf,
  data.table,
  terra,
  leaflet,
  patchwork
)

usgs_ests <- fread(here("mapping data/EPest_county_estimates_2013_2017_v2.txt"))
usgs_gly <- usgs_ests %>% filter(COMPOUND=="GLYPHOSATE")
usgs_gly_2013 <- usgs_gly %>% filter(YEAR==2013)
usgs_gly_2014 <- usgs_gly %>% filter(YEAR==2014)
usgs_gly_2015 <- usgs_gly %>% filter(YEAR==2015)
usgs_gly_2016 <- usgs_gly %>% filter(YEAR==2016)
usgs_gly_2017 <- usgs_gly %>% filter(YEAR==2017)

glimpse(usgs_gly)
dim(usgs_gly)

us_county_2013 <- st_read(here("mapping data/cb_2013_us_county_20m/cb_2013_us_county_20m.shp"))
glimpse(us_county_2013)
dim(us_county_2013)

prism_county_2013 <- readRDS(here("mapping data/weighted_area_raster_fips_tmean_daily_2013.rds"))
glimpse(prism_county_2013)
dim(prism_county_2013)

prism_county_2014 <- readRDS(here("mapping data/weighted_area_raster_fips_tmean_daily_2014.rds"))

prism_county_2015 <- readRDS(here("mapping data/weighted_area_raster_fips_tmean_daily_2015.rds"))

prism_county_2016 <- readRDS(here("mapping data/weighted_area_raster_fips_tmean_daily_2016.rds"))

prism_county_2017 <- readRDS(here("mapping data/weighted_area_raster_fips_tmean_daily_2017.rds"))

usgs_gly_2013 <- usgs_gly_2013 %>% mutate(STATE_FIPS_CODE =
                              str_pad(as.character(usgs_gly_2013$STATE_FIPS_CODE),
                              2, pad = "0"),
                COUNTY_FIPS_CODE =                  
                  str_pad(as.character(usgs_gly_2013$COUNTY_FIPS_CODE),
                              3, pad = "0"),
                fips = as.factor(paste0(STATE_FIPS_CODE, COUNTY_FIPS_CODE)))

usgs_gly_2014 <- usgs_gly_2014 %>% mutate(STATE_FIPS_CODE =
                              str_pad(as.character(usgs_gly_2014$STATE_FIPS_CODE),
                              2, pad = "0"),
                COUNTY_FIPS_CODE =                  
                  str_pad(as.character(usgs_gly_2014$COUNTY_FIPS_CODE),
                              3, pad = "0"),
                fips = as.factor(paste0(STATE_FIPS_CODE, COUNTY_FIPS_CODE)))

usgs_gly_2015 <- usgs_gly_2015 %>% mutate(STATE_FIPS_CODE =
                              str_pad(as.character(usgs_gly_2015$STATE_FIPS_CODE),
                              2, pad = "0"),
                COUNTY_FIPS_CODE =                  
                  str_pad(as.character(usgs_gly_2015$COUNTY_FIPS_CODE),
                              3, pad = "0"),
                fips = as.factor(paste0(STATE_FIPS_CODE, COUNTY_FIPS_CODE)))

usgs_gly_2016 <- usgs_gly_2016 %>% mutate(STATE_FIPS_CODE =
                              str_pad(as.character(usgs_gly_2016$STATE_FIPS_CODE),
                              2, pad = "0"),
                COUNTY_FIPS_CODE =                  
                  str_pad(as.character(usgs_gly_2016$COUNTY_FIPS_CODE),
                              3, pad = "0"),
                fips = as.factor(paste0(STATE_FIPS_CODE, COUNTY_FIPS_CODE)))

usgs_gly_2017 <- usgs_gly_2017 %>% mutate(STATE_FIPS_CODE =
                              str_pad(as.character(usgs_gly_2017$STATE_FIPS_CODE),
                              2, pad = "0"),
                COUNTY_FIPS_CODE =                  
                  str_pad(as.character(usgs_gly_2017$COUNTY_FIPS_CODE),
                              3, pad = "0"),
                fips = as.factor(paste0(STATE_FIPS_CODE, COUNTY_FIPS_CODE)))

us_county_2013 <- us_county_2013 %>% mutate(fips = 
                                  as.factor(paste0(STATEFP, COUNTYFP)))

#2013
prism_county_2013 <- prism_county_2013 %>% mutate(date=as.Date(date,
                                              format="%d/%m/%Y")) %>%
  select(-c(day,month, year))

group_by(prism_county_2013, fips) %>% tally()
#got 3108 fips codes

prism_01001 <- prism_county_2013 %>% filter(fips=="01001")
quantile(prism_01001$tmean, probs = 0.85)
prism_12033 <- prism_county_2013 %>% filter(fips=="12033")
quantile(prism_12033$tmean, probs = 0.85)

prism_county_2013_85s <- prism_county_2013 %>% group_by(fips) %>% mutate(tmean85_fullyear = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  filter(between(date, as.Date('2013-04-30'), as.Date('2013-09-01'))) %>%
  mutate(tmean85_spray = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  summarise(tmean85_fullyear=mean(tmean85_fullyear),
            tmean85_spray=mean(tmean85_spray))

#2014
prism_county_2014 <- prism_county_2014 %>% mutate(date=as.Date(date,
                                              format="%d/%m/%Y")) %>%
  select(-c(day,month, year))

group_by(prism_county_2014, fips) %>% tally()
#got 3108 fips codes

prism_01001 <- prism_county_2014 %>% filter(fips=="01001")
quantile(prism_01001$tmean, probs = 0.85)
prism_12033 <- prism_county_2014 %>% filter(fips=="12033")
quantile(prism_12033$tmean, probs = 0.85)

prism_county_2014_85s <- prism_county_2014 %>% group_by(fips) %>% mutate(tmean85_fullyear = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  filter(between(date, as.Date('2014-04-30'), as.Date('2014-09-01'))) %>%
  mutate(tmean85_spray = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  summarise(tmean85_fullyear=mean(tmean85_fullyear),
            tmean85_spray=mean(tmean85_spray))

#2015
prism_county_2015 <- prism_county_2015 %>% mutate(date=as.Date(date,
                                              format="%d/%m/%Y")) %>%
  select(-c(day,month, year))

group_by(prism_county_2015, fips) %>% tally()
#got 3108 fips codes

prism_county_2015_85s <- prism_county_2015 %>% group_by(fips) %>% mutate(tmean85_fullyear = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  filter(between(date, as.Date('2015-04-30'), as.Date('2015-09-01'))) %>%
  mutate(tmean85_spray = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  summarise(tmean85_fullyear=mean(tmean85_fullyear),
            tmean85_spray=mean(tmean85_spray))

#2016
prism_county_2016 <- prism_county_2016 %>% mutate(date=as.Date(date,
                                              format="%d/%m/%Y")) %>%
  select(-c(day,month, year))

group_by(prism_county_2016, fips) %>% tally()
#got 3108 fips codes

prism_county_2016_85s <- prism_county_2016 %>% group_by(fips) %>% mutate(tmean85_fullyear = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  filter(between(date, as.Date('2016-04-30'), as.Date('2016-09-01'))) %>%
  mutate(tmean85_spray = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  summarise(tmean85_fullyear=mean(tmean85_fullyear),
            tmean85_spray=mean(tmean85_spray))

#2017
prism_county_2017 <- prism_county_2017 %>% mutate(date=as.Date(date,
                                              format="%d/%m/%Y")) %>%
  select(-c(day,month, year))

group_by(prism_county_2017, fips) %>% tally()
#got 3108 fips codes

prism_county_2017_85s <- prism_county_2017 %>% group_by(fips) %>% mutate(tmean85_fullyear = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  filter(between(date, as.Date('2017-04-30'), as.Date('2017-09-01'))) %>%
  mutate(tmean85_spray = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  summarise(tmean85_fullyear=mean(tmean85_fullyear),
            tmean85_spray=mean(tmean85_spray))

merged_2013 <- left_join(us_county_2013, prism_county_2013_85s) %>% 
  left_join(usgs_gly_2013) %>% filter(STATEFP != "02" & STATEFP != "15" &
                                   STATEFP != "72") %>%
  mutate(gly_per_land_low = EPEST_LOW_KG/(ALAND/1e+6),
         gly_per_land_high = EPEST_HIGH_KG/(ALAND/1e+6),
         gly_per_totalarea_low = EPEST_LOW_KG/((ALAND+AWATER)/1e+6),
         gly_per_totalarea_high = EPEST_HIGH_KG/((ALAND+AWATER)/1e+6))

merged_2014 <- left_join(us_county_2013, prism_county_2014_85s) %>% 
  left_join(usgs_gly_2014) %>% filter(STATEFP != "02" & STATEFP != "15" &
                                   STATEFP != "72") %>%
  mutate(gly_per_land_low = EPEST_LOW_KG/(ALAND/1e+6),
         gly_per_land_high = EPEST_HIGH_KG/(ALAND/1e+6),
         gly_per_totalarea_low = EPEST_LOW_KG/((ALAND+AWATER)/1e+6),
         gly_per_totalarea_high = EPEST_HIGH_KG/((ALAND+AWATER)/1e+6))

merged_2015 <- left_join(us_county_2013, prism_county_2015_85s) %>% 
  left_join(usgs_gly_2015) %>% filter(STATEFP != "02" & STATEFP != "15" &
                                   STATEFP != "72") %>%
  mutate(gly_per_land_low = EPEST_LOW_KG/(ALAND/1e+6),
         gly_per_land_high = EPEST_HIGH_KG/(ALAND/1e+6),
         gly_per_totalarea_low = EPEST_LOW_KG/((ALAND+AWATER)/1e+6),
         gly_per_totalarea_high = EPEST_HIGH_KG/((ALAND+AWATER)/1e+6))

merged_2016 <- left_join(us_county_2013, prism_county_2016_85s) %>% 
  left_join(usgs_gly_2016) %>% filter(STATEFP != "02" & STATEFP != "15" &
                                   STATEFP != "72") %>%
  mutate(gly_per_land_low = EPEST_LOW_KG/(ALAND/1e+6),
         gly_per_land_high = EPEST_HIGH_KG/(ALAND/1e+6),
         gly_per_totalarea_low = EPEST_LOW_KG/((ALAND+AWATER)/1e+6),
         gly_per_totalarea_high = EPEST_HIGH_KG/((ALAND+AWATER)/1e+6))

merged_2017 <- left_join(us_county_2013, prism_county_2017_85s) %>% 
  left_join(usgs_gly_2017) %>% filter(STATEFP != "02" & STATEFP != "15" &
                                   STATEFP != "72") %>%
  mutate(gly_per_land_low = EPEST_LOW_KG/(ALAND/1e+6),
         gly_per_land_high = EPEST_HIGH_KG/(ALAND/1e+6),
         gly_per_totalarea_low = EPEST_LOW_KG/((ALAND+AWATER)/1e+6),
         gly_per_totalarea_high = EPEST_HIGH_KG/((ALAND+AWATER)/1e+6))
#cleaned merged data so that it's just for conterminous U.S. (not AK which is 02, HI which is 15, or PR which is )

gly_us_low <- ggplot() +
  geom_sf(data = merged_2013, aes(fill = EPEST_LOW_KG), linetype=0) +
  ggtitle("2013 Glyphosate per County (Low Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

gly_us_high <- ggplot() +
  geom_sf(data = merged_2013, aes(fill = EPEST_HIGH_KG), linetype=0) +
  ggtitle("2013 Glyphosate per County (High Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

gly_us_low_larea <- ggplot() +
  geom_sf(data = merged_2013, aes(fill = gly_per_land_low), linetype=0) +
  ggtitle("2013 Glyphosate per Land Area (Low Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg/km^2)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

gly_us_high_larea <- ggplot() +
  geom_sf(data = merged_2013, aes(fill = gly_per_land_high), linetype=0) +
  ggtitle("2013 Glyphosate per Land Area (High Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg/km^2)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")


temp85_plot <- ggplot() +
  geom_sf(data = merged_2013, aes(fill = tmean85_spray), linetype=0) +
  ggtitle("85th Percentile Mean Temp. by County (2013)", 
          subtitle="Conterminous U.S.") +  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Temperature (C)",
                       colors = rev(met.brewer("Hiroshige")),
                       na.value = "grey50")

temp85_plot_f_2013 <- ggplot() +
  geom_sf(data = merged_2013, aes(fill = (tmean85_spray*9/5+32)), linetype=0) +
  ggtitle("85th Percentile Mean Temp. by County (2013)", 
          subtitle="Conterminous U.S.") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Temperature (F)",
                       colors = rev(met.brewer("Hiroshige")),
                       na.value = "grey50")

temp85_plot <- ggplot() +
  geom_sf(data = merged_2017, aes(fill = tmean85_spray), linetype=0) +
  ggtitle("85th Percentile Mean Temp. by County (2017)", 
          subtitle="Conterminous U.S.") +  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Temperature (C)",
                       colors = rev(met.brewer("Hiroshige")),
                       na.value = "grey50")

temp85_plot_f_2017 <- ggplot() +
  geom_sf(data = merged_2017, aes(fill = (tmean85_spray*9/5+32)), linetype=0) +
  ggtitle("85th Percentile Mean Temp. by County (2017)", 
          subtitle="Conterminous U.S.") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Temperature (F)",
                       colors = rev(met.brewer("Hiroshige")),
                       na.value = "grey50")

hot_merged_2013 <- merged_2013 %>% filter(tmean85_spray>=26.66)
hot_plot <- ggplot() +
  geom_sf(data = hot_merged_2013, aes(fill = gly_per_land_low), linetype=0) +
  ggtitle("Glyphosate per Land Area", 
          subtitle="Hot Counties") +  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

hot_leaflet <- hot_merged_2013 %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Voyager") %>%
    addPolygons(
             stroke = FALSE,
             smoothFactor = 0,
             fillOpacity = 0.5,
             fillColor= ~
               colorNumeric("YlOrRd",gly_per_land_low)(gly_per_land_low)) %>%
  addLegend(pal = colorNumeric("YlOrRd", hot_merged_2013$gly_per_land_low),  # Define color palette
        values = hot_merged_2013$gly_per_land_low,  # Values to map
        title = "Glyphosate Usage (kg/km^2)",  # Legend title
        position = "bottomright",  # Position of the legend
        na.label = "No data"
    )
```

## Question I am working on/problem I am trying to solve

- Characterizing glyphosate levels in U.S. population 
- Relationship with environmental factors like temperature
- What areas of the U.S. have both high heat and high glyphosate usage?

## Dataset source/description

- USGS Estimated Annual Agricultural Pesticide Use for Counties of the Conterminous United States, 2013-2017 (ver. 2.0, May 2020) data were obtained from here: https://www.sciencebase.gov/catalog/item/5e95c12282ce172707f2524e

- Cartographic Boundary Files from the U.S Census Bureau were obtained from here for the year 2013: https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.2013.html#list-tab-1556094155

- PRISM daily mean temperature data for 2013 was obtained from Robbie Parks' Github: https://github.com/rmp15/PRISM-grids-into-FIPS-ZIP-censustract-USA/tree/main/output/fips/tmean

## Planned methods

- Mapping county-level glyphosate usage
- Mapping 85th percentile of county-level mean daily temperatures
- Filtering temperatures where 85th percentile mean daily temperature is 80 or higher
- Mapping glyphosate usage for just those hot counties

## Unexpected problems & how I solved them

- Glyphosate was just total amounts, not corrected for land area! So I mutated the variable to divide glyphosate by area (resulting in kg of glyphosate/km^2).

- I realized 85th percentile mean daily temp over the whole year was not as relevant as temperatures during spray season, so I had to calculate for just the spray season (approximately 1 May to 1 September). HOWEVER, spraying season will vary by growing zone and crop, so for my final analysis I need to tailor this much more.

## Glyphosate Usage in U.S.
```{r gly-use}
gly_us_low / gly_us_low_larea
```

## 85th Percentile Temperatures
```{r temps}
temp85_plot_f_2013 / temp85_plot_f_2017
```

## Glyphosate Usage in the Hottest Counties

- Used leaflet to create a zoomable map!
```{r hot-leaflet}
hot_leaflet
```
