---
title: "DaSEH Project"
output: html_document
---

# DaSEH Project

```{r clear-env, include=FALSE}
# Clear environment
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
           detach, character.only=TRUE, unload=TRUE, force=TRUE))
}
```

```{r load-install-pkg,include=FALSE}
# Load packages, installing if needed
if (!requireNamespace("pacman", quietly = TRUE))
  install.packages("pacman")
pacman::p_load(
  RColorBrewer,
  tidyr,
  tidyverse,
  ggplot2,
  MetBrewer,
  stringr,
  ggpubr,
  here,
  sf,
  data.table,
  terra,
  leaflet
)
```

## Data Used

I am using USGS pesticide estimate data, Census Bureau data on U.S. county boundaries, and historical heat records in the U.S. from PRISM. 

USGS Estimated Annual Agricultural Pesticide Use for Counties of the Conterminous United States, 2013-2017 (ver. 2.0, May 2020) data were obtained from here: https://www.sciencebase.gov/catalog/item/5e95c12282ce172707f2524e. Cartographic Boundary Files from the U.S Census Bureau were obtained from here for the year 2013: https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.2013.html#list-tab-1556094155. PRISM daily mean temperature data for 2013 was obtained from Robbie Parks' Github (https://github.com/rmp15/PRISM-grids-into-FIPS-ZIP-censustract-USA/tree/main/output/fips/tmean). Citation for this work is Tuholske, C., Lynch, V.D., Spriggs, R. et al. Hazardous heat exposure among incarcerated people in the United States. Nat Sustain 7, 394–398 (2024). https://doi.org/10.1038/s41893-024-01293-y. Data dictionaries can be found at the above links.

```{r load-raw-data}
usgs_ests <- fread(here("mapping data/EPest_county_estimates_2013_2017_v2.txt"))
usgs_gly <- usgs_ests %>% filter(COMPOUND=="GLYPHOSATE")
glimpse(usgs_gly)
dim(usgs_gly)

us_county_2013 <- st_read(here("mapping data/cb_2013_us_county_20m/cb_2013_us_county_20m.shp"))
glimpse(us_county_2013)
dim(us_county_2013)

prism_county_2013 <- readRDS(here("mapping data/weighted_area_raster_fips_tmean_daily_2013.rds"))
glimpse(prism_county_2013)
dim(prism_county_2013)

##TO-DO: LOAD ALL 2018 DATA, COMPARE WITH 2013 DATA
```

The USGS glyphosate data has 15314 rows and 6 columns. The US County shapefile data has 3221 rows and 10 columns. The PRISM county-level data for 2013 has 1134420 rows and 6 columns. These data are connected by FIPS code, which is comprised of a two-digit state code and then a three digit county code. The USGS pesticide estimate data contains a low and high estimate (EPEST_LOW_KG and EPEST_HIGH_KG).

# Data Cleaning/Wrangling

Perform any data subsetting, cleaning, or manipulation methods that were described in this course on your data. Examples are: renaming the columns, recoding values, reshaping the data, filtering the data etc. 

Create combined FIPS code (two-digit state plus three-digit county in format 01001).

```{r create-fips-codes}
usgs_gly <- usgs_gly %>% mutate(STATE_FIPS_CODE =
                              str_pad(as.character(usgs_gly$STATE_FIPS_CODE),
                              2, pad = "0"),
                COUNTY_FIPS_CODE =                  
                  str_pad(as.character(usgs_gly$COUNTY_FIPS_CODE),
                              3, pad = "0"),
                fips = as.factor(paste0(STATE_FIPS_CODE, COUNTY_FIPS_CODE)))

us_county_2013 <- us_county_2013 %>% mutate(fips = 
                                  as.factor(paste0(STATEFP, COUNTYFP)))
```

``` {r date-temp-clean}
prism_county_2013 <- prism_county_2013 %>% mutate(date=as.Date(date,
                                              format="%d/%m/%Y")) %>%
  select(-c(day,month, year))

group_by(prism_county_2013, fips) %>% tally()
#got 3108 fips codes

prism_01001 <- prism_county_2013 %>% filter(fips=="01001")
quantile(prism_01001$tmean, probs = 0.85)
prism_12033 <- prism_county_2013 %>% filter(fips=="12033")
quantile(prism_12033$tmean, probs = 0.85)


prism_county_2013_85s <- prism_county_2013 %>% group_by(fips) %>% mutate(tmean85_fullyear = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  filter(between(date, as.Date('2013-04-30'), as.Date('2022-09-01'))) %>%
  mutate(tmean85_spray = quantile(tmean, probs=0.85, na.rm=TRUE)) %>% 
  summarise(tmean85_fullyear=mean(tmean85_fullyear),
            tmean85_spray=mean(tmean85_spray))
```

```{r join-data-calc-area}
merged <- left_join(us_county_2013, prism_county_2013_85s) %>% 
  left_join(usgs_gly) %>% filter(STATEFP != "02" & STATEFP != "15" &
                                   STATEFP != "72") %>%
  mutate(gly_per_land_low = EPEST_LOW_KG/(ALAND/1e+6),
         gly_per_land_high = EPEST_HIGH_KG/(ALAND/1e+6),
         gly_per_totalarea_low = EPEST_LOW_KG/((ALAND+AWATER)/1e+6),
         gly_per_totalarea_high = EPEST_HIGH_KG/((ALAND+AWATER)/1e+6))

#cleaned merged data so that it's just for conterminous U.S. (not AK which is 02, HI which is 15, or PR which is )
```

# Data Visualization

Create some visualizations of your data using the esquisse app or the ggplot2 package.

```{r plot-gly, figures-side, fig.show="hold", out.width="50%"}
gly_us_low <- ggplot() +
  geom_sf(data = merged, aes(fill = EPEST_LOW_KG), linetype=0) +
  ggtitle("Glyphosate per County (Low Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

gly_us_high <- ggplot() +
  geom_sf(data = merged, aes(fill = EPEST_HIGH_KG), linetype=0) +
  ggtitle("Glyphosate per County (High Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

plot(gly_us_low)
plot(gly_us_high)

gly_us_low_larea <- ggplot() +
  geom_sf(data = merged, aes(fill = gly_per_land_low), linetype=0) +
  ggtitle("Glyphosate per Land Area (Low Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg/km^2)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

gly_us_high_larea <- ggplot() +
  geom_sf(data = merged, aes(fill = gly_per_land_high), linetype=0) +
  ggtitle("Glyphosate per Land Area (High Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg/km^2)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

plot(gly_us_low_larea)
plot(gly_us_high_larea)

gly_us_low_tarea <- ggplot() +
  geom_sf(data = merged, aes(fill = gly_per_totalarea_low), linetype=0) +
  ggtitle("Glyphosate per Total Area (Low Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg/km^2)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")

gly_us_high_tarea <- ggplot() +
  geom_sf(data = merged, aes(fill = gly_per_totalarea_high), linetype=0) +
  ggtitle("Glyphosate per Total Area (High Est.)") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage \n(kg/km^2)",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")
plot(gly_us_low_tarea)
plot(gly_us_high_tarea)
```

I created plots for temperature during spray season and during the full year.
``` {r plot-temp, figures-side, fig.show="hold", out.width="50%"}
temp85_plot <- ggplot() +
  geom_sf(data = merged, aes(fill = tmean85_spray), linetype=0) +
  ggtitle("85th Percentile Mean Temp. by County", 
          subtitle="Conterminous U.S.") +  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Temperature (C)",
                       colors = rev(met.brewer("Hiroshige")),
                       na.value = "grey50")

temp85_plot_f <- ggplot() +
  geom_sf(data = merged, aes(fill = (tmean85_spray*9/5+32)), linetype=0) +
  ggtitle("85th Percentile Mean Temp. by County", 
          subtitle="Conterminous U.S.") +
  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Temperature (F)",
                       colors = rev(met.brewer("Hiroshige")),
                       na.value = "grey50")
plot(temp85_plot)
plot(temp85_plot_f)
```

# Data Analysis

Perform a **simple analysis** of your data. This can involve summarizing the data to describe aspects about it (quartiles, means, range etc.) or a simple statistical test. 

Don't forget to describe what analysis you performed and why. Provide some simple **interpretation** about what your analysis might indicate about your data.

```{r spatial-analysis}
#find areas that have 85th percentile mean daily temp of 80 degrees F or higher
hot_merged <- merged %>% filter(tmean85_spray>=26.66)
hot_plot <- ggplot() +
  geom_sf(data = hot_merged, aes(fill = gly_per_land_low), linetype=0) +
  ggtitle("Glyphosate per Land Area", 
          subtitle="Hot Counties") +  theme_void(base_size = 14) +
  theme(
    plot.title = element_text(hjust = 0.5),
    plot.subtitle = element_text(colour="gray25", size=14,hjust = 0.5),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    rect = element_blank()
  ) +
  scale_fill_gradientn("Glyphosate usage",
                       colors = met.brewer("Tam"),
                       na.value = "grey50")
plot(hot_plot)

hot_leaflet <- hot_merged %>%
    st_transform(crs = "+init=epsg:4326") %>%
    leaflet(width = "100%") %>%
    addProviderTiles(provider = "CartoDB.Voyager") %>%
    addPolygons(
             stroke = FALSE,
             smoothFactor = 0,
             fillOpacity = 0.2,
             fillColor= ~
               colorNumeric("YlOrRd",gly_per_land_low)(gly_per_land_low)) %>%
  addLegend(pal = colorNumeric("YlOrRd", hot_merged$gly_per_land_low),  # Define color palette
        values = hot_merged$gly_per_land_low,  # Values to map
        title = "Glyphosate Usage (kg/km^2)",  # Legend title
        position = "bottomright"  # Position of the legend
    )
hot_leaflet
```

# Please include additional sections as desired.

```{r}

```


# Versions

Include some information about the version of R and the packages you are using by running sessionInfo().

```{r}
sessionInfo()
```

